---
title:    "ISE 5103 Intelligent Data Analytics"
subtitle: "Homework 5 - Modeling"
author:   "Daniel Carpenter & Sonaxy Mohanty"
date:     "October 2022"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    highlight: arrow
  # github_document:
  #   toc: yes
  #   toc_depth: 2
urlcolor: blue
cache: true
fig.width: 7
fig.height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

## Packages

```{r error=FALSE, message=FALSE, warning=FALSE}
# Data Wrangling
library(tidyverse)

# Modeling

# Aesthetics
library(knitr)
library(cowplot)  # multiple ggplots on one plot with plot_grid()
library(scales)
library(kableExtra)
```

## General Data Prep

### Read Data
```{r}
housingData <- read.csv('housingData.csv')
```

### Clean Numeric Data

Make dataset of `numeric` variables
```{r}
housingNumeric <- housingData %>%
  
  #selecting all the numeric data
  dplyr::select_if(is.numeric) %>%
  
  #converting the dataframe to tibble
  as_tibble()
```

Make dataset of `character` variables
```{r}
housingFactor <- housingData %>%
  
  #selecting all the numeric data
  dplyr::select_if(is.character) %>%
  
  #converting the dataframe to tibble
  as_tibble()
```


For each column with missing data, impute missing values with `PMM`

* Done with function `imputeWithPMM()` function  
* Applys function via `dplyr` logic  
* Note `seeImputation()` function to visualize the imputation from prior homework 4, not shown for simplicity in viewing


Create function to impute via `PMM`

```{r}
imputeWithPMM <- function(colWithMissingData) {
  
  # Using the mice package
  suppressMessages(library(mice))
  
  # Discover the missing rows
  isMissing <- is.na(colWithMissingData) 
  
  # Create data frame to pass to PMM inputation function from mic package
  df <- data.frame(x       = rexp(length(colWithMissingData)), # meaningless x to help show varation 
                   y       = colWithMissingData, 
                   missing = isMissing)
  
  # imputation by PMM
  df[isMissing, "y"] <- mice.impute.pmm( df$y, 
                                        !df$missing, 
                                         df$x)
  
  return(df$y)
}
```


```{r, include=FALSE}
# Note `seeImputation()` function to visualize the imputation from prior homework 4
seeImputation <- function(df, df.meanInputed, 
                          imputationMethod) {
  
  # Min/Max ranges so actual and imputed histograms align
  yMin = quantile(df.meanInputed$y, 0.05)
  yMax = max(df.meanInputed$y)
  
  # Non Altered data -------------------------------------------------
  
  meanVal = mean(df$y, na.rm=T) # mean of the non altered data
  
  # Create the plot
  p1 <- df %>%
    ggplot(aes(x = y)) +
    
    # Histogram
    geom_histogram(color = 'grey65', fill = 'grey95') +
    
    # The mean value line
    geom_vline(xintercept = meanVal, color = 'tomato3') +
    
    # Text associated with mean value
    annotate("text", 
             label = "Mean Value", 
             x = meanVal, y = 100, 
             size = 5, colour = "tomato3" ) +
    
    # Labels
    labs(title = 'Data with Missing Values',
         y     = 'Frequency', 
         x     = '' ) +
    
    xlim(yMin, yMax) + # min and max range of x axis (for equal comparison)
    theme_minimal() # Theme
  
  
  # Imputed data -------------------------------------------------

  meanValImpute = mean(df.meanInputed$y, na.rm=T)
  
  # Create the plot
  p2 <- df.meanInputed %>%
    ggplot(aes(x = y)) +
    
    # Histogram
    geom_histogram(color = 'grey65', fill = 'grey95') +
    
    # The mean value line
    geom_vline(xintercept = meanVal, color = 'tomato3') +
    
    # Text associated with mean value
    annotate("text", 
             label = "Mean Value", 
             x = meanValImpute, y = 100, 
             size = 5, colour = "tomato3" ) +
    
    # Labels
    labs(title = 'Data without Missing Values',
             subtitle = 'Using PMM',
             y = 'Frequency', 
             x = imputationMethod) +
    
    xlim(yMin, yMax) + # min and max range of x axis (for equal comparison)
    theme_minimal() # Theme
  
  # Variation scatter ----------------------------------------------------------
  
  p3 <- df.meanInputed %>% ggplot(aes(x=rexp(length(y)), y=y, color=is.na(df$y))) + 
    
    # Add points
    geom_point(alpha = 0.5) +
    
    # Colors, limits, labels, and themes
    scale_color_manual(values = c('grey80', 'tomato3'),
                       labels = c('Actuals', 'Imputed') ) +
    ylim(0, quantile(df.meanInputed$y, 0.99)) + # lower 99% of dist
    labs(title   = 'Variation of Actuals vs. Imputed Data',
         x       = 'x', 
         y       = imputationMethod,
         caption =paste0('\nUsing housing.csv data',
                         '\nOnly showing lower 99% of distribution for viewing') 
         ) +
    theme_minimal() + theme(legend.position = 'bottom',
                            legend.title    = element_blank())
  
  
  # Combine the plots for the final returned output
  combinedPlots <- plot_grid(p1, p2, p3, 
                             ncol = 1, label_size = 12,
                             rel_heights = c(1, 1.1, 1.75))
  return(combinedPlots)
}
```


Apply `PMM` function to numeric data containing null values 
```{r, warning=FALSE, message=FALSE}
# Data to store imputed values with PMM method
housingDataImputed <- housingData

# Which columns has NA's?
colNamesWithNulls <- colnames(housingNumeric[ , colSums(is.na(housingNumeric)) != 0])
colNamesWithNulls

numberOfColsWithNulls = length(colNamesWithNulls)


# For each of the numeric columns with null values
for (colWithNullsNum in 1:numberOfColsWithNulls) {
  
  # The name of the column with null values
  nameOfThisColumn <- colNamesWithNulls[colWithNullsNum]
  
  # Get the actual data of the column with nulls
  colWithNulls <- housingData[, nameOfThisColumn]
  
  # Impute the missing values with PMM
  imputedValues <- imputeWithPMM(colWithNulls)
  
  # Now store the data in the original new frame
  housingDataImputed[, nameOfThisColumn] <- imputedValues
  
  
  # Save a visualization of the imputation
  pmmVisual <- seeImputation(data.frame(y = colWithNulls),
                             data.frame(y = imputedValues),
                             nameOfThisColumn )
  
  fileToSave = paste0('OutputPMM/Imputation_With_PMM_', nameOfThisColumn, '.pdf')
  print(paste0('For imputation results of ', nameOfThisColumn, ', see ', fileToSave))
  ggsave(pmmVisual, filename = fileToSave )
}
```

\newpage

### Factor level collapse data over 5 categories
```{r}


```


# `1 (d)`

## `PCR` 

### Perform PCA analysis to see how Principal components explain variance

### Now, Apply predictions with `PCR`


\newpage

## `SVR`


\newpage

## `MARS`
