---
title:    "ISE 5103 Intelligent Data Analytics"
subtitle: "Homework 8 - Clustering"
author:   "Daniel Carpenter"
date:     "December 2022"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    highlight: arrow
    latex_engine: xelatex
  # github_document:
  #   toc: yes
  #   toc_depth: 2
urlcolor: blue
cache: true
fig.width: 7
fig.height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, 
                      warning=FALSE, 
                      echo=FALSE
                      )
```

\newpage

```{r}
# Packages

## Data Wrangling
library(tidyverse)
library(skimr)
library(lubridate) # dates

## Aesthetics
library(knitr)
library(scales)
library(kableExtra)

## Data set 
library(datasets) # chick weights

## Clustering
library(cluster)  # provides more cluster algorithms than base R (e.g. PAM)
library(useful)   # provides a plot function for clusters and "FitKMeans" and "PlotHartigan" functions
library(NbClust)  # provides tons of tools for identifying the "right" number of clusters
library(rgl)      # for 3D rotating plots
```

# General Data Prep
> For general data preparation, please see conceptual steps below. See `.rmd` file for detailed code.

## Read Training Data
Clean data to ensure each read variable has the correct data type (factor, numeric, Date, etc.)
```{r}
data("ChickWeight")
df.base <- ChickWeight; rm(ChickWeight)
```

## Create `numeric` and `factor` *base* `data frames`

Make data set of `numeric` variables called `df.base.numeric`
```{r}
df.base.numeric <- df.base %>%

  # selecting all the numeric data
  dplyr::select_if(is.numeric) %>%

  # converting the data frame to tibble
  as_tibble()
```

Make data set of `factor` variables called `df.base.factor`
```{r}
df.base.factor <- df.base %>%

  #selecting all the numeric data
  dplyr::select_if(is.factor) %>%

  #converting the data frame to tibble
  as_tibble()
```


# Data Understanding
> Create a data quality report of `numeric` and `factor` data  
> Created function called `dataQualityReport()` to create factor and numeric QA report

```{r}
# Function for data report
dataQualityReport <- function(df) {
  
  # Function to remove any columns with NA
  removeColsWithNA <- function(df) {
    return( df[ , colSums(is.na(df)) == 0] )
  }
  
  # Create Comprehensive data report using skimr package
  # This is done a bit piece-wise because PDF latex does not like the skimr package
  # Very much. So Instead of printing `skim(df)`, I have to pull the contents manually
  # Unfortunately. This is not an issue with html typically.
  dataReport <- skim(df) %>%
    rename_all(~str_replace(.,"skim_","")) %>%
    arrange(type, desc(complete_rate) ) # sort data 
  
  # Filter to the class types
  dataReport.numeric <- dataReport %>% filter(type == 'numeric') # numeric data
  dataReport.factor  <- dataReport %>% filter(type == 'factor' ) # factor  data
  
  # Remove columns that do not apply to this type of data -----------------------
  
  # numeric data
  dataReport.numeric <- removeColsWithNA(dataReport.numeric)  %>%
    
    # Clean column names by removing numeric prefix, 
    rename_all(~str_replace(.,"numeric.","")) 
    
  # factor  data
  dataReport.factor  <- removeColsWithNA(dataReport.factor ) %>%
  
    # Clean column names by removing factor  prefix
    rename_all(~str_replace(.,"factor.",""))  
  
  
  # Set up options for Display the reports
  options(skimr_strip_metadata = FALSE)
  options(digits=2)
  options(scipen=99)
  
  # Numeric report <- Get summary of data frame --------------------------------
  
    # data frame stats
    dfStats.num <- data.frame(Num_Numeric_Variables = ncol(df %>% select_if(is.numeric)),
                              Total_Observations    = nrow(df) )
    
    # Now see individual column statistics
    dfColStats.num <- dataReport.numeric %>% 
      dplyr::select(-type, -hist)
    
  
  # Factor report <- Get summary of data frame --------------------------------
  
    # Get summary of data frame
    dfStats.factor <- data.frame(Num_Factor_Variables = ncol(df %>% select_if(is.factor)),
                                 Total_Observations   = nrow(df) )
    
    # Now see individual column statistics
    dfColStats.factor <- dataReport.factor  %>% 
      dplyr::select(-type, -ordered) 
    
    
  # Return the data frames
  return(list('dfStats.num'       = dfStats.num,    
              'dfColStats.num'    = dfColStats.num,
              'dfStats.factor'    = dfStats.factor, 
              'dfColStats.factor' = dfColStats.factor))
}
```

## Numeric Data Quality Report

```{r}
# Get the factor and numeric reports
initialReport <- dataQualityReport(df.base)

# Numeric data frame stats
initialReport$dfStats.num %>% kable()

# Numeric column stats
initialReport$dfColStats.num %>%
  kable() %>% kable_styling(font_size=7, latex_options = 'HOLD_position') # numeric data
```

<!-- \newpage -->

## Factor Data Quality Report

```{r}
# factor data frame stats
initialReport$dfStats.factor %>% kable()

# factor column stats
initialReport$dfColStats.factor %>%
  kable() %>% kable_styling(font_size=7, latex_options = 'HOLD_position') # numeric data
```


## Review Target Cluster

```{r}
df.base %>%
  ggplot(aes(y     = weight,
             x     = Time,
             color = Diet
             )) +

  # The points  
  geom_point() +
  
  # Direction of the trend.
  # Generally the higher the diet, the larger the weight
  # Diet 3 is exceptional to that statement
  geom_smooth(fill = 'grey85') +
  
  # Labels
  labs(title    = 'How Experimental Diets Affect Chick Weights (Nominal Data)',
       subtitle = 'Note Adjusted for time since chick birthed',
       x        = 'Number of Days since Chick Birthed',
       y        = 'Weight of the Chick (gm)',
       caption  = '\nGrouped by individual chick on a given day since birthed') + 
  
  # Colors and aesthetics
  scale_color_brewer(palette = 'Set2') +
  theme_minimal()
```

\newpage

## Clustering Analysis

```{r}
# Elbow plot function provided in lecture
#create our own plot function to look for "Within cluster Sum of square error 'elbow' plot"
#defaults to 15 as clusters max

wssplot <- function(data, nc=15){                    

  par(mfrow=c(1,2))
  
  wss <- NULL  
  pctExp <-NULL
  
  for (k in 1:nc)
  {
     kclus <- kmeans(data, centers=k)
     wss[k] <- kclus$tot.withinss      #store the total within SSE for given k
     pctExp[k] <- 1-wss[k]/kclus$totss
  }
  
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")

  plot(1:nc, pctExp, type="b", xlab="Number of Clusters",
       ylab="Pct Explained")
  
  par(mfrow=c(1,1))
}
```

```{r}
# Data prep
df <- df.base # copy to produce altered data (liking scaling)

set.seed(5013) # for similar reproducing

# center and scale dat
df[, c('weight', 'Time')] <- scale( df[, c('weight', 'Time')] )
```

```{r}
# Clustering analysis - identify number of clusters

# Check for "elbow" using lecture method
wssplot(df[, 1:3], nc=10)

# Now try nbClust to see how kmeans identifies it
df.temp <- df
df.temp$Chick <- as.numeric(df$Chick) # needs to be numeric, not factor. so temp change

# Find the suggested number of clusters
NbClust(df.temp[, 1:3], method="kmeans")
```

```{r}
# Actual cluster - using the above number of suggested clusters 
# Most suggested between (3-5), so try to 4

NUM_CLUSTERS = 4

cl <- kmeans(df[,1:3], NUM_CLUSTERS, nstart=100)
df$kMeansCluster <- as.factor(cl$cluster)
# head(df)

# Check the clustering if it lined up
# A few line up, but not necessariliy ordered well
table( df$Diet, df$kMeansCluster )

df %>%
  ggplot(aes(y     = weight,
             x     = Time,
             color = kMeansCluster
             )) +

  # The points  
  geom_point() +
  
  # Direction of the trend.
  # Generally the higher the diet, the larger the weight
  # Diet 3 is exceptional to that statement
  geom_smooth(fill = 'grey85') +
  
  # Labels
  labs(title    = 'How Experimental Diets Affect Chick Weights (K-Means Clustered Data)',
       subtitle = 'Note Adjusted for time since chick birthed',
       x        = 'Number of Days since Chick Birthed',
       y        = 'Weight of the Chick (gm)',
       caption  = '\nGrouped by individual chick on a given day since birthed') + 
  
  # Colors and aesthetics
  scale_color_brewer(palette = 'Set2') +
  theme_minimal()
```



```{r}
# hiearchial clustering 

# hiearchial clustering requires a distance matrix
di <- dist(df, 1:3, method="euclidean")   # with hiearchical clustering, only need distance matrix

hc.dward <- hclust(di, method="ward.D")

df$hierarchicalClusterWard <- as.factor(cutree(hc.dward, k=NUM_CLUSTERS))   #cutting at k=3, here are the assignments

# quick check to see how our 'natural clusters' align with Diet data
# Does a GREAT job!
table( df$Diet, df$hierarchicalClusterWard )

df %>%
  ggplot(aes(y     = weight,
             x     = Time,
             color = hierarchicalClusterWard
             )) +

  # The points  
  geom_point() +
  
  # Direction of the trend.
  # Generally the higher the diet, the larger the weight
  # Diet 3 is exceptional to that statement
  geom_smooth(fill = 'grey85') +
  
  # Labels
  labs(title    = 'How Experimental Diets Affect Chick Weights (Hierarchical Clustered Data - Ward)',
       subtitle = 'Note Adjusted for time since chick birthed',
       x        = 'Number of Days since Chick Birthed',
       y        = 'Weight of the Chick (gm)',
       caption  = '\nGrouped by individual chick on a given day since birthed') + 
  
  # Colors and aesthetics
  scale_color_brewer(palette = 'Set2') +
  theme_minimal()
```


```{r}
# Try daisy method

d <- daisy(df[, 1:3], stand=TRUE)

hc.avg<-hclust(d, method="average")  # notice how balanced the clusters are
hc.avg<-hclust(d, method="complete")  # notice how balanced the clusters are
hc.avg<-hclust(d, method="ward.D")  # notice how balanced the clusters are
plot(hclus)


df$hierarchicalClusterAvg <- as.factor(cutree(hc.avg, k=NUM_CLUSTERS))   #cutting at k=4, here are the assignments

table( df$Diet, df$hierarchicalClusterAvg )


df %>%
  ggplot(aes(y     = weight,
             x     = Time,
             color = hierarchicalClusterAvg
             )) +

  # The points  
  geom_point() +
  
  # Direction of the trend.
  # Generally the higher the diet, the larger the weight
  # Diet 3 is exceptional to that statement
  geom_smooth(fill = 'grey85') +
  
  # Labels
  labs(title    = 'How Experimental Diets Affect Chick Weights (Hierarchical Clustered Data - Average)',
       subtitle = 'Note Adjusted for time since chick birthed',
       x        = 'Number of Days since Chick Birthed',
       y        = 'Weight of the Chick (gm)',
       caption  = '\nGrouped by individual chick on a given day since birthed') + 
  
  # Colors and aesthetics
  scale_color_brewer(palette = 'Set2') +
  theme_minimal()
```
