---
title: "Homework 1"
author: "Daniel Carpenter"
date: "August 2022"
format: 
  pdf:
    toc: true
    toc-depth: 1
    number-sections: true
    # mainfont: Arial
    highlight-style: github #arrow
    # code-block-bg: true
    code-block-border-left: "#D6D6D6"

editor: visual
---

\newpage

## Using R: Vectors

### (a)

Create a vector with 10 numbers (3, 12, 6, -5, 0, 8, 15, 1, -10, 7) and assign it to x.

```{r}
x <- c(3, 12, 6, -5, 0, 8, 15, 1, -10, 7)
```

### (b)

Using the commands seq, min, and max with one line of code create a new vector y with 10 elements ranging from the minimum value of x to the maximum value of x.

```{r}
# Min, max of x vector
xMin = min(x)
xMax = max(x)
INCREMENT = (xMax - xMin) / (length(x) - 1) # size of each step

# Create the vector from min of x to max of x while maintaining length of 10
y = seq(min(x), max(x), INCREMENT)
y # display

# Prove is 10 elements:
length(y)
```

### (c)

Compute the sum, mean, standard deviation, variance, mean absolute deviation, quartiles, and quintiles for x and y.

```{r}
```

### (d)

Use sample() to create a new 7 element vector z by using R to randomly sample from x with replacement.

```{r}
```

### (e)

Use t.test() to compute a statistical test for differences in means between the vectors x and y. Are the differences in means significant?

```{r}
```

### (f)

To sort a data frame in R, use the order() function. Sort the vector x and re-run the t-test as a paired t-test.

```{r}
```

### (g)

Create a logical vector that identifies which numbers in x are negative.

```{r}
```

### (h)

Use this logical vector to remove all entries with negative numbers from x. (Make sure to overwrite the vector x so that the new vector x has 8 elements!)

```{r}
```
